ESRC SDAI Application

### 2019-06-18 case_for_support.tex
Anna update

\subsection*{1. Who is abused and who reports abuse?}

 \textcolor{red}{\textit{Short intro about aims.}}
In our first research project, we will be using the CSEW and WMP datasets to answer two main research questions. First, we will identify the demographic and socio-economic characteristics of the victims of domestic abuse. Second, we will estimate the extent of underreporting, and explore the characteristics of victims who are least likely to report domestic abuse. Exploring the characteristics of groups vulnerable to domestic abuse, the true extent of the problem of underreporting and the factors affecting the likelihood of reporting is crucial for designing effective interventions and prevent potential cases of domestic homicide.

\textcolor{red}{\textit{What we know so far}} from the \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/articles/domesticabusefindingsfromthecrimesurveyforenglandandwales/yearendingmarch2018}{CSEW} : 7.9\% women and 4.2\% men reported to have experienced domestic abuse in the past year; women between 20-24 are most likely to have experienced domestic abuse; for men, age doesn't represent a significant risk factor;  divorced or separated status are significant risk factors across both sexes; long-term illness or disability is a significant predictor across both sexes; single parent households are the most vulnerable; women in the lowest household income bracket are the most likely to have experienced domestic abuse, the same is true for men but the pattern is less pronounced; only 17\% of people who have experienced domestic abuse in the past year have reported to the police, there is no significant difference in this proportion across genders.

\textcolor{red}{\textit{Detailed plan, highlighting added value}}

@@

Second, we can validate some of these findings using the WMP data. This dataset contains information about the sex and age of victims. We can contrast the sex-age victimisation numbers from the WMP with corresponding population estimates to validate the findings of the CSEW\textit{(we were initially talking about using victims of burglary, but maybe this is better?)}. We will also use the WMP data on the offender's sex and age to characterise domestic abuse perpetrators.

@@

\textbf{Aim 2.}  While the strength of the CSEW is that it provides reliable estimates of the proportion of the population who have been a victim of a certain crime, it cannot capture the true extent of domestic abuse victimisation for several reasons. First, because it does not collect information on repeat domestic abuse victimisation (an inherent characteristic of domestic abuse cases). Second, because it only focuses on certain sub-categories of domestic abuse (physical violence, sexual assault, stalking), while ignores others (criminal damage, verbal abuse, fraud, burglary, theft). Third, it does not collect information on whether different offences against the same victim were perpetrated by the same offender. In contrast, the WMP data allows us to tell if a reported case was domestic abuse related, regardless of the actual offence committed. We can estimate the co-morbidity of self-reported victimisation by crime type using the CSEW. We can contrast this with the estimated co-morbidity of reported victimisation from the police data. \textit{is this silly? then we can maybe estimate what percentage of true victimisation is domestic-abuse related?}

@@

\textbf{Aim 3.} Police forces in England, including the West Midlands Police, have previously been criticised for mis-recording violent cases of violent crime, including \href{https://www.bbc.co.uk/news/uk-england-468676578}{domestic abuse.} Mis-recording these already heavily underreported crimes decrease victim's trust in the Criminal Justice System and their willingness to report subsequent incidents. By contrasting the extrapolated estimates from the CSEW (which is representative sample) with WMP data on the true number of reported cases, we can estimate the true extent of the mis-recording of domestic abuse cases by the WMP.

@@

%We will also extend the analysis to provide a more general description of the victims of abuse by considering

@@

\textit{Core outcomes}: A demographic description of victims, a demographic description of perpetrators, and an estimate of the reporting gap, and a demographic description of who does not report domestic abuse.

>>>>

\subsection*{1. Understanding domestic abuse: victim characteristics, extent of domestic abuse and police mis-recording}

 \textcolor{red}{\textit{Short intro about aims, highlighting added value.}}
In our first research project, we will be using the CSEW and WMP datasets to investigate three main research questions. First, we will use the CSEW and WMP to identify the demographic and socio-economic characteristics of the victims of domestic abuse, and the WMP to characterise perpetrators of reported cases. In addition, we aim to explore the predictors of the decision to report and the level of harm suffered. Exploring the characteristics of groups vulnerable to domestic abuse, the true extent of the problem of underreporting and the factors affecting the likelihood of reporting is crucial for designing effective interventions and prevent potential cases of domestic homicide. Second, we will estimate the true extent of domestic abuse victimisation. Given the multifaceted nature of domestic abuse and the problem of underreporting, it is very hard to reliably estimate the extent of the problem and the societal costs of domestic abuse. We wish to do so by combining the strengths of the two datasets (that is, the CSEW provides a more reliable measure of the true rate of victimisation across different offence types, while the WMP provides unique information on whether various types of offences were domestic abuse related). Finally, we will estimate the extent of police mis-recording domestic abuse. Improving the police's response to reported incidents is key in encouraging victims to report domestic abuse. By contrasting estimated levels from reported incidents within the West Midlands from the two datasets, we can identify any reporting gaps between the two sources of data that might arise from police mis-recording.

\textcolor{red}{\textit{What we know so far}}

Aim 1. \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/articles/domesticabusefindingsfromthecrimesurveyforenglandandwales/yearendingmarch2018}{CSEW} : 7.9\% women and 4.2\% men reported to have experienced domestic abuse in the past year; women between 20-24 are most likely to have experienced domestic abuse; for men, age doesn't represent a significant risk factor;  divorced or separated status are significant risk factors across both sexes; long-term illness or disability is a significant predictor across both sexes; single parent households are the most vulnerable; women in the lowest household income bracket are the most likely to have experienced domestic abuse, the same is true for men but the pattern is less pronounced; only 17\% of people who have experienced domestic abuse in the past year have reported to the police, there is no significant difference in this proportion across genders.

Aim 2. The prevalence of DA is usually based on the CSEW, so I think the second analysis can add value.

Aim 3. \href{https://www.justiceinspectorates.gov.uk/hmicfrs/news/news-feed/little-evidence-of-improvements-in-crime-recording-at-west-midlands-police/}{HMICFRS}

\textcolor{red}{\textit{Detailed plan}}

@@

Second, we can validate some of these findings using the WMP data. This dataset contains information about the sex and age of victims. We can contrast the sex-age victimisation numbers from the WMP with corresponding population estimates to validate the findings of the CSEW\textit{(we were initially talking about using victims of burglary, but maybe this is better?)}. We will also use the WMP data on the offender's sex and age to provide a characterisation of domestic abuse perpetrators.

@@

\textbf{Aim 2.}  While a unique strength of the CSEW is that it provides reliable estimates of the proportion of the population who have been a victim of a certain crime, it cannot capture the true extent of domestic abuse victimisation for several reasons. First, because it does not collect detailed information on repeat domestic abuse victimisation (an inherent characteristic of domestic abuse cases), but only on the number of incidents involving physical violence perpetrated by household members. In addition, victims who suffer frequent, ongoing abuse are less likely to accurately remember the exact number of incidents. Second, because it only focuses on certain sub-categories of domestic abuse (physical violence, sexual assault, stalking), while ignores others (criminal damage, verbal abuse, financial abuse, burglary). In addition, it does not collect information on whether different offences against the same victim were perpetrated by the same offender. In contrast, the WMP data allows us to tell if a reported case was domestic abuse related, regardless of the actual offence committed. In addition, it allows us to identify various incidents where the same victim-offender pair were involved. We can estimate the co-morbidity of self-reported victimisation by crime type using the CSEW. We can contrast this with the estimated co-morbidity of reported victimisation from the police data. \textit{is this silly? then we can maybe estimate what percentage of true victimisation is domestic-abuse related? the £66 billion cost estimate is based on the CSEW}

@@

\textbf{Aim 3.} Police forces in England, including the West Midlands Police, have previously been criticised for mis-recording violent cases of violent crime, including \href{https://www.bbc.co.uk/news/uk-england-468676578}{domestic abuse.} Mis-recording these already heavily underreported crimes decrease victim's trust in the Criminal Justice System and their willingness to report subsequent incidents. By contrasting the extrapolated estimates from the CSEW (which is representative sample) with WMP data on the true number of reported cases, we can estimate the true extent of the mis-recording of domestic abuse cases by the WMP. Given the repeated cross-sectional nature of the CSEW, and the longitudinal nature of the WMP, we can quantify the proportion of mis-recorded incidents over time.

@@

@@

\textcolor{red}{\textit{Core outcomes}} 
A deeper understanding of the socio-economic predictors of domestic abuse victimisation, the decision to report and the level of harm suffered. The first estimation of the true extent of domestic abuse-related victimisation across all offence types. An estimation of the extent of police (WMP) mis-recording of domestic abuse over time.
